Mining Scenario/Transaction Pairs HW/SW Co-verification Using Symbolic Execution

### 2017-04-02 paper.tex
contribution

\newcommand{\tool}[1]{\textsc{#1}\xspace}
\newcommand{\symex}{\tool{path-symex}}
\newcommand{\ebmc}{\tool{ebmc}}
\newcommand{\hector}{\tool{hector}}
\newcommand{\slec}{\tool{slec}}
\newcommand{\symexv}{\tool{path-symex 5.0}}
\newcommand{\cbmcv}{\tool{cbmc}}
\newcommand{\cbmcver}{\tool{cbmc 5.2}}
\newcommand{\ebmcv}{\tool{ebmc 4.2}}
\newcommand{\hwcbmcv}{\tool{hw-cbmc}}
\newcommand{\hwcbmcver}{\tool{hw-cbmc 5.0}}
\newcommand{\acdcl}{\tool{acdcl}}
\newcommand{\summarizer}{\tool{summarizer 1.0}}
\newcommand{\abc}{\tool{ABC}}
\newcommand{\yosys}{\tool{Yosys 0.5}}
\newcommand{\verifoxver}{\tool{VerifOx 0.1}}
\newcommand{\verifox}{\tool{VerifOx}}
\newcommand{\iimc}{\tool{iimc}}
\newcommand{\klee}{\tool{klee}}
\newcommand{\symbiotic}{\tool{symbiotic}}
\newcommand{\kleever}{\tool{klee 1.0.0}}
\newcommand{\symbioticver}{\tool{symbiotic 3.0.1}}

@@

A firmware only exercise some fragment of hardware state-space in a typical hardware/software co-design~\cite{dac17,taa,ibm-other}.  However, there is no automatic hardware/software co-verification tool that can exploit this interaction behavior.  For building effective and scalable formal co-verification tool, it is therefore imperative to automatically mine only relevant scenario/transaction pairs that occurs in practical co-designs.  In this paper, we attempt to bridge this gap by presenting a technique that automatically mine relevant HW/SW interaction behavior using symbolic execution.
A firmware services the requests from
the application software layer or its interacting hardware~\footnotesize{Note that whenever we refer to ``hardware'' from this point onwards, we mean
an RTL implementation and not a TLM.}  components. 
Similarly, the hardware services the requests from the firmware or the
physical environment.  Thus, the firmware and its interacting hardware
components can be seen as a set of services.  In this paper, we call a
hardware RTL service a {\em Hardware Transaction} $(\mathcal{HT})$.  Closest
to our notion of $\mathcal{HT}$ is the micro-architecture-level transaction
by Mahajan et al.~\cite{mcbmq}.  We call a FW service a {\em Firmware
Transaction} ($\mathcal{FT}$).  $\mathcal{FT}$ is an untimed finite-state
machine that is manifested as control-flow paths in the FW, similar to
transactions in TLM~\cite{codes14}.  A~{\em scenario} $(\mathcal{S})$ in a
co-design typically configures the co-verification environment in a way that
restricts the interaction between $\mathcal{FT}$ and $\mathcal{HT}$.  In
other words, given a scenario $\mathcal{S}$, the FW only exercises a fragment
of the HW state-space.  This renders many interactions between
$\mathcal{HT}$ and $\mathcal{FT}$ infeasible. For example, a scenario in \emph{Universal Asynchronous Receiver Transmitter} (UART) IP can configure the UART in \emph{transmitter} mode which restricts the communication between the FW and HW to set of services that only transfers data from the FW to the serial output port of the transmitter HW while completely bypassing the receiver HW.    A {\em Scenario/Transaction}
pair is a set of valid interactions between $\mathcal{HT}$ and
$\mathcal{FT}$ for a particular scenario~$\mathcal{S}$.  
It is necessary to automatically mine Scenario/Transaction pairs for scalable and effective co-verification.

@@

%===============================================================================
%\subsection{Contributions}
%===============================================================================
In this paper, we make the following contributions:
\begin{enumerate}
\item Translating Hardware/Software Co-designs to an unified 
sequential software model, represented in C.  
verification using Software Model Checkers.
\item {\em Co-verification using Symbolic Execution:} 
In this paper, we introduce a notion of {\em transaction} 
in HW RTL and FW models. We present a unified FW-HW model, 
called {\em Program Netlist}, in C, for the co-verification task. 
To this end, we present a path-based symbolic execution algorithm that is 
able to automatically find scenario/transaction pairs from a 
program netlist and check their validity. 
To the best of our knowledge, this is the first attempt to 
perform path-based symbolic execution of 
C (FW) and Verilog RTL (HW) models by synthesizing 
them to a unified program netlist model.

\item {\em {\verifox} -- A tool for HW/SW co-verification:} We address the
co-verification problem at post-RTL phase and present a tool, \verifox, for
bounded HW/SW co-verification of IP level and SoC level design.  \verifox
supports IEEE 1364-2005 System Verilog standards and the C89, C99 standards. 
\verifox also supports SAT and SMT backends for constraint solving.  We
distribute the HW/SW co-verification tools, \hwcbmcv and \verifox, along
with hardware benchmarks in Verilog RTL and firmware C models
here~\footnote{https://drive.google.com/folderview?id=0BynC6sd-wTYgWVQ4bXNtRDB5a2c\&usp=sharing}. 
\end{enumerate}

@@

\Omit{
Even though \verifox supports SMT solvers such as
Z3~\cite{z32008} as backend solvers, the current implementation communicates
with these solvers using files rather than APIs.  Since a new solver
instance is generated for every query, incremental encoding cannot be used. 
Due to this, the runtimes of \verifox with SMT solvers on benchmarks are
significantly higher.  As the current implementation does not support
incremental encoding with SMT solvers, the runtimes would not provide a fair
comparison of \verifox with SMT versus \verifox with SAT solvers. 
Therefore, we are omitting the results obtained using SMT solvers.

>>>>

\newcommand{\tool}[1]{\textsc{#1}\xspace}
\newcommand{\symex}{\tool{path-symex}}
\newcommand{\ebmc}{\tool{ebmc}}
\newcommand{\hector}{\tool{hector}}
\newcommand{\slec}{\tool{slec}}
\newcommand{\symexv}{\tool{path-symex 5.0}}
\newcommand{\cbmcv}{\tool{cbmc}}
\newcommand{\cbmcver}{\tool{cbmc 5.2}}
\newcommand{\ebmcv}{\tool{ebmc 4.2}}
\newcommand{\hwcbmcv}{\tool{hw-cbmc}}
\newcommand{\hwcbmcver}{\tool{hw-cbmc 5.0}}
\newcommand{\acdcl}{\tool{acdcl}}
\newcommand{\summarizer}{\tool{summarizer 1.0}}
\newcommand{\abc}{\tool{ABC}}
\newcommand{\yosys}{\tool{Yosys 0.5}}
\newcommand{\verifoxver}{\tool{VerifOx 0.1}}
\newcommand{\verifox}{\tool{CoVeriIf}}
\newcommand{\iimc}{\tool{iimc}}
\newcommand{\klee}{\tool{klee}}
\newcommand{\symbiotic}{\tool{symbiotic}}
\newcommand{\kleever}{\tool{klee 1.0.0}}
\newcommand{\symbioticver}{\tool{symbiotic 3.0.1}}

@@

A firmware only exercise some fragment of hardware state-space in a typical 
hardware/software co-design~\cite{dac17,taa,ibm-other}.  However, there is no 
automatic hardware/software co-verification tool that can exploit this interaction 
behavior.  For building effective and scalable formal co-verification tool, it is 
therefore imperative to automatically mine only relevant scenario/transaction pairs 
that occurs in practical co-designs.  In this paper, we attempt to bridge this gap by 
presenting a technique that automatically mine relevant HW/SW interaction behavior 
using symbolic execution.  A firmware services the requests from
the application software layer or its interacting hardware~\footnote{Note that whenever 
we refer to ``hardware'' from this point onwards, we mean an RTL implementation 
and not a TLM.}  components. 
Similarly, the hardware services the requests from the firmware or the
physical environment.  Thus, the firmware and its interacting hardware
components can be seen as a set of services.  In this paper, we call a
hardware RTL service a {\em Hardware Transaction} $(\mathcal{HT})$.  Closest
to our notion of $\mathcal{HT}$ is the micro-architecture-level transaction
by Mahajan et al.~\cite{mcbmq}.  We call a FW service a {\em Firmware
Transaction} ($\mathcal{FT}$).  $\mathcal{FT}$ is an untimed finite-state
machine that is manifested as control-flow paths in the FW, similar to
transactions in TLM~\cite{codes14}.  A~{\em scenario} $(\mathcal{S})$ in a
co-design typically configures the co-verification environment in a way that
restricts the interaction between $\mathcal{FT}$ and $\mathcal{HT}$.  In
other words, given a scenario $\mathcal{S}$, the FW only exercises a fragment
of the HW state-space.  This renders many interactions between
$\mathcal{HT}$ and $\mathcal{FT}$ infeasible. For example, a scenario in \emph{Universal Asynchronous Receiver Transmitter} (UART) IP can configure the UART in \emph{transmitter} mode which restricts the communication between the FW and HW to set of services that only transfers data from the FW to the serial output port of the transmitter HW while completely bypassing the receiver HW.    A {\em Scenario/Transaction}
pair is a set of valid interactions between $\mathcal{HT}$ and
$\mathcal{FT}$ for a particular scenario~$\mathcal{S}$.  
It is necessary to automatically mine Scenario/Transaction pairs for scalable and effective co-verification.

@@

%===============================================================================
%\subsection{Contributions}
%===============================================================================
In this paper, we make the following contributions:
\begin{enumerate}

  \item We present two approaches for formal hardware/software co-verification
    -- 1) A \emph{monolithic} approach that clearly separates the symbolic
    execution of hardware and firmware model and generates a monolithic SAT
    formula for verification, 2) A \emph{path-based} approach that constructs a
    unified co-verification model by sequentializing the hardware and firmware
    models and generates multiple SAT formulas through combined symbolic
    execution of the unified model.

  \item We present an automatic HW/SW co-verifcation tool called, \hwcbmcv,
    which follows the monolithic verification technique.  Our path-based 
    approach is implemented in a tool called, \verifox. Both tools can handle
    co-designs where the hardware is implemented in Verilog RTL and firmware is  
    specified in ANSI-C.  We distribute the HW/SW co-verification tools, \hwcbmcv and 
    \verifox, along with our co-design benchmarks in Verilog RTL and C models
here~\footnote{https://drive.google.com/folderview?id=0BynC6sd-wTYgWVQ4bXNtRDB5a2c\&usp=sharing}.

  \item We show that the proposed path-based approach translates the HW/SW co-verification problem
     into a purely software verification problem.  Our experiments show that the representation
     of the sequentialized unified co-verification model in software languages allows us to 
     mine relevant transaction/snecario pairs for effective co-verification.  To the 
     best of our knowledge, this is the first attempt to synthesize a pure
     software model for co-verification purpose.

   \item We present an unified property specification language in \hwcbmcv and 
     \verifox that allows designers to write meaningful formal co-specification 
     over hardware and firmware events. We believe that the formal
     co-specification model will significantly ease the adoption of formal tools 
     for co-verification of industrial strength co-designs.

\end{enumerate}

@@

\Omit{
\item {\em {\verifox} -- A tool for HW/SW co-verification:} We address the
co-verification problem at post-RTL phase and present a tool, \verifox, for
bounded HW/SW co-verification of IP level and SoC level design.  \verifox
supports IEEE 1364-2005 System Verilog standards and the C89, C99 standards. 
\verifox also supports SAT and SMT backends for constraint solving.  
}
\Omit{
Even though \verifox supports SMT solvers such as
Z3~\cite{z32008} as backend solvers, the current implementation communicates
with these solvers using files rather than APIs.  Since a new solver
instance is generated for every query, incremental encoding cannot be used. 
Due to this, the runtimes of \verifox with SMT solvers on benchmarks are
significantly higher.  As the current implementation does not support
incremental encoding with SMT solvers, the runtimes would not provide a fair
comparison of \verifox with SMT versus \verifox with SAT solvers. 
Therefore, we are omitting the results obtained using SMT solvers.
